---
title: "03_Analysis"
output:
  html_document:
    df_print: paged
---

The notebook shows the process of analysis for different metrics. The corpus comprised several author's corpora:  
  
  8 male authors
- nkrja selected authors: Karamzin, Zhukovsky, Pushkin, Baratynsky, Lermontov   
- poems published in periodicals between 1835 and 1840 by: Benediktov, Kozlov, Timofeev
  
  7 female authors + one mistification 
- poems published between 1835 and 1840 by female authors: Rostopchina, Teplova,  Shakhova, Pavlova, Kul'man;  
- poems by earlier women poets M. Pospelova and A. Bunina;  
- mysterious "female prodigy" A. Smirnova.  

```{r, echo=TRUE, comment=FALSE, include=FALSE}
library(tidyverse)
library(tidytext)

library(ggrepel)
library(wesanderson)
theme_set(theme_minimal())
```


Basic statistics and overview of the data. Some authors have unproportionally big corpora, so I randomly select 10000 words from their corpora for further analysis.

```{r}
getwd()
```


```{r}
load("../data/02_2_analysis_data_tidy.Rda")

n_poems <- analysis %>% 
  group_by(author, text_ID) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(author) %>% 
  count() %>% 
  rename(n_poems = n)

n_tokens <- analysis %>% 
  group_by(author) %>% 
  count() %>% 
  ungroup() %>% 
  rename(n_tokens = n)

auth_gender <- tibble(author = c("БаратынскийЕА", "Бенедиктов В.Г.", "Бунина А.",
                                 "ЖуковскийВА", "КарамзинНМ", "Козлов И.И.", "Кульман Е.Б.",
                                 "ЛермонтовМЮ", "Павлова К.К.", "Поспелова М.", "ПушкинАС", 
                                 "Ростопчина Е.П.", "Смирнова А.", "Теплова Н.С.", 
                                 "Тимофеев А.В.", "Шахова Е.Н." ),
                      gender = c("m", "m", "f", "m", "m", "m", "f", "m", "f", 
                                 "f", "m", "f", "f", "f", "m", "f"))

base_statistics <- auth_gender %>% 
  left_join(n_poems, by = "author") %>% 
  left_join(n_tokens, by = "author") %>% 
  arrange(gender, desc = T)

base_statistics

rm(n_poems, n_tokens, auth_gender)
```

``` {r}
glimpse(analysis)

analysis$pos %>% table()

analysis %>% 
  filter(!author %in% c("ПушкинАС", "ЖуковскийВА", "КарамзинНМ", "ЛермонтовМЮ", 
                        "Бунина А.", "БаратынскийЕА")) %>% 
  group_by(author) %>% 
  count() %>% 
  ungroup() %>% 
  summarise(mean_corpus = mean(n))
# mean number of words in a corpus from periodicals: ~7000 words

# randomly select only 10 000 words form large corpora
random_big <- analysis %>%
  filter(author %in% c("ПушкинАС", "ЖуковскийВА", "КарамзинНМ", "ЛермонтовМЮ", 
                        "Бунина А.", "БаратынскийЕА")) %>%
  group_by(author) %>% 
  sample_n(10000) %>% 
  ungroup()

randomized <- analysis %>% 
  filter(!author %in% c("ПушкинАС", "ЖуковскийВА", "КарамзинНМ", "ЛермонтовМЮ", 
                        "Бунина А.", "БаратынскийЕА")) %>% 
  rbind(random_big) 

#write.csv(randomized, "../data/02_2_random_sample_1.csv")
```
```{r, echo = FALSE}
# randomized <- read.csv("../data/02_2_random_sample_1.csv")
```

```{r}
# example for calculating proportion of POS for each author
nouns <- randomized %>% 
  filter(pos == "S") %>% 
  group_by(author) %>% 
  summarise(n_nouns = n(),
            n_f_nouns = sum(fem),
            prop_f_nouns = n_f_nouns/n_nouns) %>% 
  arrange(prop_f_nouns, desc = FALSE)

nouns
```
```{r, echo = FALSE}
print("Proportion of feminine POS")

adj <- randomized %>% 
  filter(pos == "A") %>% 
  group_by(author) %>% 
  summarise(n_adj = n(),
            n_f_adj = sum(fem),
            prop_f_adj = n_f_adj/n_adj)

# verbs & participles 
verbs <- randomized %>% 
  filter(pos == "V") %>% 
  group_by(author) %>% 
  summarise(n_verb = n(),
            n_f_verb = sum(fem),
            prop_f_verb = n_f_verb/n_verb)

# APRO - adjectival pronouns
# SPRO - pronouns
apro <- randomized %>% 
  filter(pos == "APRO") %>% 
  group_by(author) %>% 
  summarise(n_apro = n(), 
            n_f_apro = sum(fem), 
            prop_f_apro = n_f_apro/n_apro)

spro <- randomized %>% 
  filter(pos == "SPRO") %>% 
  group_by(author) %>% 
  summarise(n_spro = n(), 
            n_f_spro = sum(fem), 
            prop_f_spro = n_f_spro/n_spro)

intj_temp <- randomized %>% 
  filter(pos == "INTJ") %>% 
  group_by(author) %>% 
  count() 

intj <- tibble(
  author = base_statistics$author,
  n_tokens = base_statistics$n_tokens) %>% 
  left_join(intj_temp, by = "author") %>% 
  group_by(author) %>% 
  summarise(
    n_intj = n,
    prop_intj = n/n_tokens)

morphology <- left_join(
  base_statistics, nouns %>% select(author, prop_f_nouns), by = "author") %>% 
  left_join(adj %>% select(author, prop_f_adj), by = "author") %>% 
  left_join(verbs %>% select(author, prop_f_verb), by = "author") %>% 
  left_join(apro %>% select(author, prop_f_apro), by = "author") %>% 
  left_join(spro %>% select(author, prop_f_spro), by = "author") %>% 
  left_join(intj %>% select(author, prop_intj), by = "author")

morphology
# write.csv(morphology, "data/morphology_random_results.csv")
```

```{r, echo = F}
means_random <- randomized %>% 
  filter(pos %in% c("S", "V", "A", "APRO", "SPRO")) %>% 
  group_by(pos) %>%
  summarise(n = n(),
            f_pos = sum(fem),
            proc_f = f_pos / n)

total_tokens <- randomized %>%
  count() # total tokens

proc_int <- (randomized %>% filter(pos == "INTJ") %>% count()) / total_tokens[1,1]

#proc_int

print("Mean values for parts of speech in feminine forms")
means_random
```
```{r, include = F, echo = F}
# randomized <- read.csv("data/random_sample_1.csv")
```

```{r}
randomized <- randomized %>% 
  mutate(lemma = str_extract(analysis, "\\[\\{'lex': '.*?',")) %>% 
  mutate(lemma = str_remove_all(lemma, "[[:punct:]]|lex|[[:space:]]")) 
```
``` {r, echo = FALSE}
print("extract lemmas from analysis")
randomized %>% 
  select(word, lemma) %>% 
  head()
```

NB! These dictionaries for abstractness / concreteness indexes were prepared by a research group from Kazan Federal University; current lists are publicly available and can be found at: https://kpfu.ru/tehnologiya-sozdaniya-semanticheskih-elektronnyh.html  
On methodology of lists compilation see: Solovyev V.D., Volskaya Y.A., Andreeva M.I., Zaikin A.A. (2022). "Russian dictionary with concreteness/abstractness indices." Russian Journal for Linguistics, 26-2, 515-549.  
  
```{r}
# Abstractness / concreteness dictionaries cleaning & merge
dict_adj <- read.delim("../data/dicts/adj_concreteness.csv", sep = ";") %>% 
  mutate(word = str_remove_all(word, "[[:punct:]]|[[:space:]]"), 
         index_c_a = str_replace_all(index_c_a, ",", "\\.")) %>% 
  mutate(index_c_a = as.numeric(index_c_a))
dict_noun <- read.delim("../data/dicts/nouns_concreteness.csv", sep = ";") %>% 
  mutate(word = str_remove_all(word, "[[:punct:]]|[[:space:]]"), 
         index_c_a = str_replace_all(index_c_a, ",", "\\.")) %>% 
  mutate(index_c_a = as.numeric(index_c_a))

glimpse(dict_adj)
glimpse(dict_noun)

dictionaries <- rbind(dict_adj, dict_noun) %>% rename(lemma = word) %>% select(-score)

randomized <- randomized %>% left_join(dictionaries, by = "lemma")
```

```{r, echo = FALSE}
head(randomized)

randomized %>% 
  select(lemma, index_c_a) %>% 
  filter(!is.na(index_c_a))
```

```{r}
glimpse(randomized)

abstract_concrete <- randomized %>% 
  filter(!is.na(index_c_a)) %>% 
  group_by(author) %>% 
  summarise(mean_ac = mean(index_c_a))

head(abstract_concrete, 16)
```

Add vector model lists of closely thematic words
```{r}
w2v_lists <- read.csv("../data/02_1_notions_vectors.csv")
head(w2v_lists)
```
How percentage of a theme-related words is calculated:
```{r}
n_tokens <- randomized %>% 
  group_by(author) %>% 
  count(sort = T) %>% 
  rename(total = n) %>% 
  arrange(author, desc = T)

x <- as.vector(str_split_fixed(w2v_lists$words[1], pattern = " ", n = 30))

# example how it works
randomized %>% 
  filter(lemma %in% x) %>% 
  group_by(author) %>% 
  count() %>% 
  arrange(author, desc = T) %>% 
  left_join(n_tokens, by = "author") %>% 
  mutate(proc_theme = n/total) %>% 
  rename(!!x[1]:=proc_theme)
```

```{r, warning=FALSE}
# main counter
x <- NULL
y <- NULL
themes <- NULL
w2v_results <- list()

for (i in 1:nrow(w2v_lists)) {
  # select list of words
  x <- as.vector(str_split_fixed(w2v_lists$words[i], pattern = " ", n = 30))
  
  # save theme name
  themes[i] <- x[1]
  
  # count number of words related
  y <- randomized %>% 
  filter(lemma %in% x) %>% 
  group_by(author) %>% 
  count() %>% 
  arrange(author, desc = T) %>% 
  left_join(n_tokens, by = "author") %>% 
  mutate(proc_theme = n/total)
  
  # store results
  w2v_results[[i]] <- y$proc_theme
}

w2v_results
#themes

# table pivoting
w2v_results_t <- as.tibble(do.call(rbind, w2v_results))
colnames(w2v_results_t) <- as.vector(n_tokens$author)

w2v_results_t$theme <- themes

w2v_results_t

w2v_fin <- w2v_results_t %>% 
  pivot_longer(!theme, names_to = "author", values_to = "proc_theme") %>% 
  pivot_wider(names_from = theme, values_from = proc_theme)

w2v_fin

# write.csv(w2v_fin, "data/w2v_perc_table.csv")
```

Merge all results together
```{r}
# glimpse(morphology)
# glimpse(abstract_concrete)
# glimpse(w2v_fin)

# x <- as.data.frame(summary(w2v_fin))
# x

main_results <- morphology %>% 
  left_join(abstract_concrete, by = "author") %>% 
  left_join(w2v_fin, by = "author") %>% 
  mutate(author = str_replace_all(author, "(\\w)(\\w)(\\w)$", "\\1 \\2\\.\\3\\."))

head(main_results)

```

```{r}
# lables
ids <- main_results$author
genders <- main_results$gender
```

```{r}
# very fast projection for everything (should be messy??)

t <- main_results %>% select(-author, -gender, -n_poems, -n_tokens)

values_matrix <- t %>% 
  as.matrix() 

distances <- dist(values_matrix, method = "euclidean") 

# distances <- wide %>%
#   as.matrix() %>% 
#   JSD(unit="log2") %>% 
#   `rownames<-`(ids) 

mds <- cmdscale(distances, eig = TRUE, k = 2)

proj <- tibble(x = mds$points[,1], 
               y = mds$points[,2],
               author = ids, # does this work the same as `rownames<-`(ids) ??
               gender = genders)
  
               
proj %>% 
  ggplot(aes(x, y, color = gender)) + 
  geom_point(data = proj, aes(x, y)) + 
  geom_text_repel(aes(label = author)) +
  theme(legend.position = "None")

```
Projection only on POS features
```{r, echo = F}
t <- morphology %>% select(-author)

values_matrix <- t %>% 
  as.matrix() 

distances <- dist(values_matrix, method = "euclidean") 

# distances <- wide %>%
#   as.matrix() %>% 
#   JSD(unit="log2") %>% 
#   `rownames<-`(ids) 

mds <- cmdscale(distances, eig = TRUE, k = 2)

proj <- tibble(x = mds$points[,1], 
               y = mds$points[,2],
               author = ids, # does this work the same as `rownames<-`(ids) ??
               gender = genders)
  
               
proj %>% 
  ggplot(aes(x, y, color = gender)) + 
  geom_point(data = proj, aes(x, y), color = wes_palette("Rushmore1")[4]) + 
  geom_label_repel(aes(label = author),
                  box.padding = 1.2, 
                  max.overlaps = Inf,
                  ) + 
  theme(legend.position = "None") + 
  labs(title = "Projection based on female-POS usage")

```
```{r, echo=FALSE, warning=FALSE}
t <- w2v_fin %>% select(-author)

values_matrix <- t %>% 
  as.matrix() 

distances <- dist(values_matrix, method = "euclidean") 

# distances <- wide %>%
#   as.matrix() %>% 
#   JSD(unit="log2") %>% 
#   `rownames<-`(ids) 

mds <- cmdscale(distances, eig = TRUE, k = 2)

proj <- tibble(x = mds$points[,1], 
               y = mds$points[,2],
               author = ids, # does this work the same as `rownames<-`(ids) ??
               gender = genders)
  
               
proj %>% 
  ggplot(aes(x, y, color = gender)) + 
  geom_point(data = proj, aes(x, y)) + 
  geom_text_repel(aes(label = author)) +
  theme(legend.position = "None") + 
  labs(title = "Projection based on the usage of thematic words")

```

#### Grouping w2v features to 5 groups
```{r, include = TRUE}
female <- rowMeans(w2v_fin[, c("любовь", "дом", "семья", "цветок", "бал")])
body <- rowMeans(w2v_fin[, c("уста", "грудь", "глаз", "лобзание")])
poetry <- rowMeans(w2v_fin[, c("поэт", "муза", "лира")])
war <- rowMeans(w2v_fin[, c("война", "русский", "страна", "служить")])
sorrow <- rowMeans(w2v_fin[, c("смерть", "печаль")])

prop_f_pron <- rowMeans(morphology[, c("prop_f_apro", "prop_f_spro")])

grouped_results <- morphology %>% 
  left_join(abstract_concrete, by = "author") %>% 
  cbind(female, body, poetry, war, sorrow, prop_f_pron) %>% 
  select(-prop_f_apro, -prop_f_spro)

glimpse(grouped_results)

grouped_results <- grouped_results %>% 
  filter(author != "Смирнова А.")

ids <- grouped_results$author
genders <- grouped_results$gender
```

```{r}
t <- grouped_results %>% select(-author, -gender, -n_poems, -n_tokens)

values_matrix <- t %>% 
  as.matrix() 

distances <- dist(values_matrix, method = "euclidean") 

# distances <- wide %>%
#   as.matrix() %>% 
#   JSD(unit="log2") %>% 
#   `rownames<-`(ids) 

mds <- cmdscale(distances, eig = TRUE, k = 2)

proj <- tibble(x = mds$points[,1], 
               y = mds$points[,2],
               author = ids, # does this work the same as `rownames<-`(ids) ??
               gender = genders)
  
               
proj %>% 
  ggplot(aes(x, y, color = gender)) + 
  geom_point(data = proj, aes(x, y)) + 
  geom_text_repel(aes(label = author)) + 
  # geom_label_repel(aes(label = author),
  #                 box.padding = 1,
  #                 #max.overlaps = Inf,
  #                 ) +
  theme(legend.position = "None") + 
  labs(title = "Projection based on the grouped features") + 
  scale_fill_manual(values = c(wes_palette("Rushmore1")[5], wes_palette("Rushmore1")[4])) +
  scale_color_manual(values = c(wes_palette("Rushmore1")[5], wes_palette("Rushmore1")[4]))
```

```{r}
# ggsave("../plots/projection.png", plot = last_plot(), width = 8, height = 6, dpi = 300, bg = "white")

# write.csv(grouped_results, "../data/02_2_grouped_results.csv")

summary(grouped_results)
```

```{r, eval = F}
# scaled table

dat <- read.csv("data/02_2_grouped_results.csv")

glimpse(dat)

dat_for_scale <- dat %>% 
  select(-X, -author, -gender, -n_poems, -n_tokens)

x <- as.data.frame(scale(dat_for_scale)) %>% mutate_if(is.numeric, round, digits = 4)

dat_scaled <- cbind(dat$author, x) 

# write.csv(dat_scaled, "data/02_02_grouped_results_scaled.csv")
```


```{r}
glimpse(grouped_results)

fem_pos <- rowMeans(grouped_results[, c("prop_f_nouns", "prop_f_adj", "prop_f_verb", "prop_f_pron")])

grouped_results %>% 
  select(author, gender, female) %>% 
  cbind(fem_pos) %>% 
  ggplot(aes(x = fem_pos, y = female, color = gender)) + 
  geom_point() + 
  geom_text_repel(aes(label = author))

```

```{r}
grouped_results %>% 
  select(author, gender, mean_ac) %>% 
  cbind(fem_pos) %>% 
  ggplot(aes(x = fem_pos, y = mean_ac, color = gender)) + 
  geom_point() + 
  geom_text_repel(aes(label = author))
```

```{r}
grouped_results %>% 
  select(author, gender, war) %>% 
  cbind(fem_pos) %>% 
  ggplot(aes(x = fem_pos, y = war, color = gender)) + 
  geom_point() + 
  geom_text_repel(aes(label = author))
```

```{r}
grouped_results %>% 
  select(author, gender, body) %>% 
  cbind(fem_pos) %>% 
  ggplot(aes(x = fem_pos, y = body, color = gender)) + 
  geom_point() + 
  geom_text_repel(aes(label = author))
```

